### 2019-09-17 paper_neural_network.Rmd
Added text to paper and added figures.

---
title: "NeuralNetworkVisualization"
output:
  pdf_document:
    toc: true
    number_sections: true
---
```{r global_options, include = FALSE}
knitr::opts_chunk$set(fig.width = 7, fig.height = 4, fig.path = 'Figs/',
                      fig.align = "center", warning = FALSE, message = FALSE)
```
# Introduction

# Theory

## Marginal Effects

## Neural Networks

## Partial Dependence Plots

## Discussion of Partial Dependence Plots

# Introducing the "NeuralNetworkVisualization" - package

## Comparison with existing packages

## Unique selling point of NeuralNetworkVisualization
\newpage

# Code structure
Our package consists of four core functionalities:

1. Fitting the model with the **NeuralNetwork** class

2. Preparing the data for visualization using the **prepare_data** function

3. Creating the partial dependence plots using the function 
**plot_partial_dependencies**

4. Create the visualizations inside a shiny app with the **run_shiny_app** 
function

Implementing a package that is able to create partial dependence plots requires 
having the ability to fit neural networks easily. **neuralnet** is the most 
common R package used for fitting neural networks but it has some minor 
inconveniences. First of all, factor response and predictor variables have to be 
expanded to a set of dummy variables. In the neuralnet package this has to be 
done manually before fitting the model using the **class.ind** function from the 
**nnet** package. Furthermore, if the range of the predictors is different, 
scaling or data normalization has to be performed. Again, in the **neuralnet** 
function this has to be done before fitting the model. Additionally, our package 
functionality has as a requirement to remember all network specifications 
provided by the user since we need to refit the model multiple times for 
creating the bootstrap confidence intervals for the partial dependence plots. 
Hence, we decided to write a wrapper class around the **neuralnet** function.

The **NeuralNetwork** class takes similar inputs as the **neuralnet** function 
and keeps track of all inputs for further model refitting for the confidence 
intervals. Furthermore, users don't need to expand factor variables to a set 
of dummy variables before fitting the model. This is being handled automatically 
inside the **NeuralNetwork** class. Besides that, a scale parameter has been 
added with which users can decide if they want the data to be normalized. 
Everything corresponding to the model fitting part has been saved in the 
**neural_net_class.R** file.

Data preparation for plotting is hidden from the end user. The **prepare_data** 
function is responsible for the plotting data creation as explained with the 
algorithm in chapter 2.3. Besides preparing the plotting data the function does 
also add the lower and upper levels for the confidence intervals if wanted by 
the user. All created visualizations use this function to get the necessary data 
for plotting. If the user decides to add a confidence interval, the bootstrap 
data will be recreated for each new plot. Moreover if the scale parameter of the
**NeuralNetwork** has been set to *TRUE*, the data will be descaled back to the 
original data range for plotting. The functionality for the data preparation is 
stored in the **plotting_data_preparation.R** file.

We decided not to add the bootstrap data creation inside the **NeuralNetwork** 
class since some users might not be interested in adding a confidence interval 
to the partial dependence plots and having them wait longer then necessary for 
their visualizations is inappropriate. If the confidence interval should be 
added, keep in mind that for each bootstrap iteration the neural network will be 
refitted with sampled data. This procedure might take some time, since fitting a
neural network depends on the available data, the number of hidden layers and 
the threshold for the minimization of the prediction error.

After having fit the model, the user can jump directly into producing partial 
dependence plots with the **plot_partial_dependencies** function. There the user 
can select all (default behavior), multiple or just a single predictor for 
which to create the partial dependence plots. Besides that, one can select the 
lower and upper quantiles for the confidence intervals, the number of bootstrap 
iterations and whether it should be a ggplot or plotly visualization. The 
**neural_net_visualization.R ** file contains the plotting function definitions.

The partial dependence plots can be created via a shiny user interface with the 
**run_shiny_app** function. There the user just has to upload the 
**NeuralNetwork** which should be stored as a **.rds** file.
\newpage

# Example
This chapter will explain how to use the R package. The examples will be based 
on the famous iris data set where the neural network will be used to predict the 
flower species based on different measurements.

First of all, lets fit the neural network using the **NeuralNetwork** class. 
Compared to the **neuralnet** package less data preprocessing has to be done by
the user since the *Species* column does not have to be expanded into dummy 
variables and scaling is not necessary since the predictors do have similar data 
ranges.

```{r, results = "hide"}
library(NeuralNetworkVisualization); library(datasets); data("iris")
model_data <- iris

set.seed(1)
model <- NeuralNetwork(
    Species ~ .,
    data = model_data, layers = c(5, 5), rep = 5, linear.output = FALSE,
    scale = FALSE, err.fct = "ce", stepmax = 1000000, threshold = 0.0001)
```

Now it is easily possible to create the 
partial dependence plots using the **plot_partial_dependencies** function. If 
the user just provides the model as input, the function will create the partial 
dependence plots for all predictors without a confidence interval.

```{r}
plot_partial_dependencies(model)
```

The partial dependence plots show how the probability of a flower being in a 
specific category change given a certain value for the predictor of interest and 
averaging out all other predictors.

Let's take a look at the partial dependence plot for *Petal.Width* and add a 
bootstrap confidence interval to capture the uncertainty of the partial 
dependence computation. The thick line will represent the average partial 
dependency for the number of bootstrap iterations and the lower and upper bound 
the respective quantiles.

```{r}
plot_partial_dependencies(model, predictors = "Petal.Width", 
                          probs = c(0.05, 0.95), nrepetitions = 5)
```

Now that the confidence interval has been added to the partial dependence plots, 
the interpretation can additionally take the uncertainty of the effect of a 
change of the predictor on the response category into account.

# Outlook